Representing and Reasoning About Spatial Regions Defined by Context

### 2012-01-20 scenarios/cdsr/trunk/papers/AAAI2012/qsr-extraction.tex
edited qsr text a bit

For each object that Dora detects we compute 8 spatial relations between that object and each of the objects adjacent to it; adjacency is determined using a voronoi diagram, as is standard in geometric reasoning \cite{Forbus/etal2003}. The relations we compute between each given \emph{landmark} object and its adjacent neighbour are analogous to the cardinal and intermediate points on the compass when the compass is centered on the object. The canonical directions of these relations are defined using the following vectors: $<0,1>$, $<1,1>$, $<1,0>$, $<1,-1>$, $<0,-1>$, $<-1,-1>$, $<-1,0>$, $<-1,1>$.

We generate these relations as follows. Taking each object in the room in turn to be the current landmark, we translate the origin of the room to the landmark's centroid. This results in the coordinates of the all the other objects in the room being translated into a frame of reference whose origin is the centroid of the landmark.  For each neighbour object adjacent to the landmark we compute the 8 relations by calculating the inner angle between the vector from the origin (the landmark's centroid) and the neighbour's location and the direction vector of the relation, as defined above. This angle is computed by taking the dot product of these two vectors. $\dots$

%%\from{Nick}{I have changed the below paragraph to match the actual numbers and images. I've marked what I changed}
%%Desk 1 and Desk 2 are not labeled in the Figure.

%Consider the situation in Figure~\ref{fig:dora-spatial} where the robot is facing two desks in a row\footnote{Please note that although the desks in front of the robot in Figure~\ref{fig:dora-spatial} appear distributed left to right from the robot's current position, they are actually distributed bottom to top, according to our definitions, in the global coordinate frame.}. Because we represent the desks as distinct points, \fw{(rcc8-DC Desk1 Desk2)} states that they are disjoint. To account for their relative positions, neither \fw{(leftOf Desk1 Desk2)} nor \fw{(leftOf Desk2 Desk1)} are true because the x coordinate of \fw{Desk1} is within a threshold of the x coordinate of \fw{Desk2}. The statement \fw{(below Desk1 Desk2)} is true because \fw{Desk1} has a lower y coordinate. Each desk is also related to the room by the non-tangential proper part relation, \fw{rcc8-NTPP}. These qualitative spatial relationships provide the structure necessary for analogical processing.

>>>>

For each object that Dora detects we compute 8 spatial relations between that object and each of the objects adjacent to it; adjacency is determined using a voronoi diagram, as is standard in geometric reasoning \cite{Forbus/etal2003}. The model used to compute these relations was inspired by the literature on modeling the semantics of spatial terms \cite{kelleher/costello:09,kelleher/vanGenabith:2006,regier/carlson:2001,gapp:1994}. The model accommodates both direction and distance as factors in the relative position of objects.

The relations we compute between each given \emph{landmark} object and its adjacent neighbours are analogous to the cardinal and intermediate points on the compass when the compass is centered on the object. The canonical directions of these relations are defined using the following vectors: $\langle0,1\rangle$, $\langle1,1\rangle$, $\langle1,0\rangle$, $\langle1,-1\rangle$, $\langle0,-1\rangle$, $\langle-1,-1\rangle$, $\langle-1,0\rangle$, $\langle-1,1\rangle$. The predicates used to denote these relations are named accordingly, e.g. \textit{xOneYZero}, \textit{xOneYOne}, \textit{xZeroYOne}, \textit{xOneYMinOne}, etc.

We generate the strengths of these spatial relations as follows. First we compute the maximum distance between any two points in the room, this value is used to normalize the distances between objects. Next, taking each object in turn to be the landmark, we translate the origin of the room to the landmark's centroid. This results in the coordinates of the all the other objects in the room being translated into a frame of reference whose origin is the centroid of the landmark.  We then compute the strength of each of the 8 spatial relations between the landmark and each of the objects adjacent to it by calculating: (a) the distance between the landmark's centroid and the adjacent object's location, and (b) the inner angle between the direction vector of the relation and the vector from the origin (the landmark's centroid) to the neighbour's location. These two spatial components are integrated to compute the strength of a given relationships using Equation~\ref{eq:spatial}. Figure \ref{fig:spr} provides a visualization of a spatial relationship across a region.

\from{Nick}{TODO: The following equation breaks the Matrix}

\begin{equation}
strength = (innerAngle > 90^{\circ}? 0 : (1- \frac{innerAngle}{90^{\circ}}) * (1- \frac{distance}{maxDistance}))
\label{eq:spatial}
\end{equation}

% These spatial relationships between adjacent objects provide the structure necessary for analogical processing.

\begin{figure}
\centerline{
\includegraphics[width=2in]{./images/spatialRelations/xZeroYOne.png}
}
\caption{A visualisation of a the strength of a spatial relation across a region. In this figure the landmark is located at the center of the image and the direction vector used was $\langle0,1\rangle$ (i.e. above of the landmark). The darker the pixel the stronger the spatial relation is deemed to be at that point.}
\label{fig:spr}
\end{figure}
